---
title: "Care homes - descriptive summaries"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = F, message = F, warning = F)

library(tidyverse)
library(lubridate)
theme_set(theme_bw())

source("./analysis/data_setup.R")

```

## Care home characteristics: summary tables

```{r tables}

# Total care homes in analysis dataset
N_ch_tot <- n_distinct(dplyr::select(ch_chars, household_id, msoa))

# Number of carehomes per MSOA
ch %>% 
  group_by(msoa) %>% 
  summarise(
    n_resid = n(),
    n_ch = n_distinct(household_id),
    n_gp = n_distinct(practice_id)) %>%
  ungroup() -> per_msoa

knitr::kable(per_msoa) %>%
  kableExtra::kable_styling() %>%
  kableExtra::scroll_box(width = "100%", height = "200px")

# Average number of carehomes per msoa
per_msoa %>%
  pull(n_ch) %>%
  summary()
```


Summarise characteristics overall, and by whether or not care home had any covid-event during period (ever-affected). Size of each home estimated as number of patients registered under that household ID, but actual capacity may be larger. 

```{r tab1}

ch_overall <- ch_chars %>%
  mutate(ever_affected = "Overall")

chars_waffect <- ch_chars %>%
  full_join(ch_first_event) %>%
  mutate(ever_affected = ifelse(ever_affected, "Affected","Unaffected")) %>%
  bind_rows(ch_overall) %>% 
  mutate(ever_affected = factor(ever_affected, 
                                levels = c("Overall","Affected","Unaffected"))) 

# Tabulate care home type (factor):
chars_waffect %>%
  group_by(ever_affected, ch_type) %>%
  summarise(n_resid = n()) %>%
  ungroup() %>%
  mutate(ch_type = replace_na(as.character(ch_type), "missing")) %>%
  pivot_wider(names_from = ch_type, values_from = n_resid, names_prefix = "Type:") %>%
  rowwise() %>%
  mutate(n_resid = sum(c_across(cols = -ever_affected))) %>% 
  ungroup() %>%
  mutate_at(vars(-n_resid, -ever_affected), 
            function(x) paste0(x, " (", round(x/.$n_resid,1), ")")) %>%
  column_to_rownames("ever_affected") %>%
  dplyr::select(-n_resid) -> tab_type

# Tabulate other variables
chars_waffect %>%
  group_by(ever_affected) %>%
  summarise(N = n(),
            `No. residents` = sum(n_resid, na.rm = T),
            ch_size_mean = round(mean(ch_size, na.rm = T),1),
            ch_size_sd = round(sqrt(var(ch_size, na.rm = T)),2),
            `% rural` = round(sum(rural_urban == "rural")/N, 2),
            imd_mean = round(mean(imd, na.rm = T)),
            imd_sd = round(sqrt(var(imd, na.rm = T)),1)) %>%
  mutate(N_perc = round(N/N_ch_tot,2),
         # `N (%)` = paste0(N, " (",N_perc,")"),
         `size mean(sd)` = paste0(ch_size_mean, " (",ch_size_sd,")"),
         `IMD mean(sd)` = paste0(imd_mean, " (",imd_sd,")")) %>%
  ungroup() %>%
  column_to_rownames("ever_affected") %>%
  dplyr::select(N, `% rural`, `IMD mean(sd)`, `size mean(sd)`, `No. residents`) %>%
  cbind(tab_type) -> tab1

knitr::kable(tab1)

```


Age, dementia status and ethnicity of care home residents, stratified by whether or not their home was affected (percentages out of total residents in that stratum):

```{r tab2}

ch_resid_all <- ch %>%
  mutate(ever_affected = "Overall")

ch %>%
  full_join(ch_first_event) %>%
  mutate(ever_affected = ifelse(ever_affected,"Affected","Unaffected")) %>%
  bind_rows(ch_resid_all) %>% 
  mutate(ever_affected = factor(ever_affected, 
                                levels = c("Overall","Affected","Unaffected"))) -> ch_resid_all
  
ch_resid_all %>%
  group_by(ever_affected) %>%
  summarise(`No. residents` = n(),
            med_age = round(median(age, na.rm = T)),
            q_age = paste(round(quantile(age, 
                                         probs = c(0.25, 0.75), 
                                         na.rm = T)), collapse = ", "),
            n_dem = sum(dementia, na.rm = T)
            ) %>% 
  mutate(`age med[IQR]` = paste0(med_age, " [",q_age,"]"),
         `dementia n(%)` = paste0(n_dem, " (",round(n_dem/`No. residents`,1),")")) %>%
  ungroup() %>%
  column_to_rownames("ever_affected") %>%
  dplyr::select(`No. residents`, `age med[IQR]`, `dementia n(%)` ) -> tab_age

ch_resid_all %>%
  group_by(ever_affected, ethnicity) %>%
  summarise(n_resid = n()) %>%
  ungroup() %>%
  mutate(ethnicity = replace_na(as.character(ethnicity), "missing")) %>%
  pivot_wider(names_from = ethnicity, values_from = n_resid, names_prefix = "Ethn:") %>%
  rowwise() %>%
  mutate(n_resid = sum(c_across(cols = -ever_affected))) %>% 
  ungroup() %>%
  mutate_at(vars(-n_resid, -ever_affected), 
            function(x) paste0(x, " (", round(x/.$n_resid,1), ")")) %>%
  column_to_rownames("ever_affected") %>%
  dplyr::select(-n_resid) -> tab_ethn

tab2 <- cbind(tab_age, tab_ethn)

knitr::kable(tab2)

```


## Care home survival

```{r ch_surv}

# Cumulative care home survival
ch_long %>%
  group_by(date) %>%
  filter(first_event > date) %>%
  summarise(n = n_distinct(household_id)) %>%
  ggplot(aes(date, n)) +
  geom_line() +
  labs(title = "Survival of care homes from COVID-19 introduction",
       x = "", y = "No. without event")

```

## Community burden

```{r community_inc}

# Average daily incidence
comm_prev %>%
  group_by(date) %>%
  summarise(probable_cases_rate = mean(probable_cases_rate)) -> comm_prev_avg

# Community incidence over time
comm_prev %>%
  ggplot(aes(date, probable_cases_rate)) +
  geom_line(aes(group = msoa), alpha = 0.1) +
  geom_line(data = comm_prev_avg, col = "white", lty = "dashed", lwd = 1.5) + 
  labs(title = "Probable cases per 100,000, by MSOA",
       x = "", y = "Rate")

```


## Community incidence versus care home introduction

```{r ch_vs_comm}

lmk_data %>%
  mutate(event_ahead = as.factor(event_ahead)) %>%
  pivot_longer(c("probable_cases_rate","probable_chg7","probable_roll7")) %>%
  ggplot(aes(event_ahead, value)) +
  geom_boxplot() +
  coord_flip() +
  facet_grid(rows = "name", scales = "free") +
  labs(title = "Community incidence versus 14-day-ahead introduction",
       y = "Daily probable cases in community, per 100,000",
       x = "Introduction in next 14 days")

```


## Hospital discharges of care home residents

```{r disch}

lmk_date %>%
  group_by(date) %>%
  summarise(n_disch = sum(n_disch, na.rm = T)) %>%
  ggplot(aes(date, n_disch)) +
  geom_line() + 
  labs(title = "Hospital discharges of care home residents",
       x = "", y = "Count")

```


## Community, care home and older population epidemics

```{r comp_epidemics}

input %>%
  filter(!is.na(primary_care_case_probable)) %>%
  mutate(group = case_when(care_home_type == "U" & age < 70 ~ "Community",
                           care_home_type != "U" ~ "Care home",
                           care_home_type == "U" & age >= 70 ~ "Community, aged 70+")) %>%
  group_by(primary_care_case_probable, group) %>%
  count() %>% 
  ggplot(aes(primary_care_case_probable, n, col = group)) +
  geom_line() +
  labs(title = "Daily probable cases identified through primary care",
       col = "Population",
       x = "Date",
       y = "Count")

```
